### 2018-06-15 source/Ch2-intro/calibration.tex
why cheesoid exist

\pg
There are three `generations' of calibration methods, of increasing complexity. We will describe them in terms of the RIME, showing how each generation increases in generality to account for more exotic effects. They are referred to interchangeably as `nth-generation calibration' or `nGC' methods.

@@

\pg
First-generation calibration methods (1GC methods) consist of open-loop calibration. This relies entirely on instrument stability, and thus imposes significant design constraints on radio telescopes. It consists of briefly observing a calibrator before and after each observation run to find a gain factor and offset error\footnote{For a concrete example, see \href{http://www.analog.com/en/analog-dialogue/articles/open-loop-calibration-techniques.html}{the open-loop calibration techniques page of analog.com}.}

@@

\pg
In this section, I will discuss the \emph{facet-based} approach taken by \citetads{2017arXiv171202078T}. The approach cannot be divorced from imaging, for the simple reason that solving for and applying Jones matrices\footnote{See \cref{chapter.RIME} for a full discussion on calibration and Jones matrices} in different directions requires image-plane knowledge of the sky brightness distribution $\Bmatrix$. % The imaging challenges this methods introduces will be discussed in [href to relevant imaging section].

>>>>

\pg
We can distinguish three `generations' of calibration methods, of increasing complexity. We will describe them in terms of the RIME, showing how each generation increases in generality to account for more exotic effects. They are referred to interchangeably as `nth-generation calibration' or `nGC' methods.

@@

\pg
First-generation calibration methods (1GC methods) consist of open-loop calibration. This relies entirely on instrument stability, and thus imposes significant design constraints on radio telescopes. It consists of briefly observing an external calibrator before and after each observation run to find a gain factor and offset error\footnote{For a concrete example, see \href{http://www.analog.com/en/analog-dialogue/articles/open-loop-calibration-techniques.html}{the open-loop calibration techniques page of analog.com}.}

@@

\pg
In this section, I will discuss the \emph{facet-based} approach taken by \citetads{2017arXiv171202078T}. The approach cannot be divorced from imaging, for the simple reason that solving for and applying Jones matrices\footnote{See \cref{section.RIME} for a full discussion on calibration and Jones matrices} in different directions requires image-plane knowledge of the sky brightness distribution $\Bmatrix$. % The imaging challenges this methods introduces will be discussed in [href to relevant imaging section].

### 2018-06-15 source/Ch2-intro/imaging.tex
why cheesoid exist

\pg
There are two quantities of interest to all astronomers: sensitivity and angular resolution\footnote{Other kinds of resolutions - in time and frequency, for example - are also extremely important to many astronomers, but are not what concern us here.}. An instrument's sensitivity is a function of its collecting area\footnote{It is also a function of technological factors, of course, but \emph{ceteris paribus}, a more sensitive telescope means a telescope with a wider collecting area}. Resolution, for well-designed instruments\footnote{By this, we mean that we assume that an instrument is also designed to optimise resolution.}, is limited by diffraction. This introduces specific issues in the radio domain. Radio waves have very long wavelengths - often comparable to meters, rather than the $\sim100-1000$nm wavelengths of optical light. Achieving a resolution comparable to those of optical telescopes would thus require making telescopes with apertures tens of millions of times larger than those already titanic instruments!

@@

\pg
Interferometric observations follow a very familiar pattern: the elements of the interferometric array (which can be single dishes or phased arrays\footnote{Described in \cref{sec.visibility}} of smaller antennas) measure a voltage. This voltage is correlated between antennas, as will be described shortly. Each of these correlations correspond to a single Fourier mode of the sky, as per the Zernike-van Cittert theorem (cf. \cref{sec.imag.psf}). They must be calibrated, by comparing the observations of a calibrator source and modelled visibilities. Calibration is described in much greater detail in \cref{sec.RIME}. Then, the calibrated visibilities are used to reconstruct an image of the sky, which often requires deconvolution due to the sparsity of interferometric arrays.

@@

\pg
Different astrophysical sources are located very far apart, and so their signals do not interfere. They are measured additively in both antennas. Provided that the signal from both sources is coherent when observed by the dishes, the correlation between the voltages measured by antennas A and B will simply be the sum of the voltage correlations associated with individual sources. This is shown in \cref{fig.visibility}: the radio signal from two different sources, at two different locations, is picked up simultaneously by both radio antennas. Because only the correlation between both antennas is measured, the signals do not interfere. % - i.e. the interferometric signal from different sources are additive.
\pg
Note that in Fig. \ref{fig.visibility.measure}, neither source is at the zenith. This introduces the notion of the \emph{effective baseline} which is shorter than the \emph{physical baseline} when measuring a signal from a source that is not at the zenith. Sources in different positions in the sky will be measured by an interferometric array, but the same array will have different effective baselines for different sources. To point an interferometric array in a specific direction in the sky, one must digitally introduce a phase delay in each antenna (or, for fundamental interferometer elements such as the LOFAR tiles or NenuFAR, by playing with the cable length between each dish and the correlator). This ensures that the constituent antennas' sensitivities are maximised in the direction of interest. The position of sources are then given in terms of the cardinal angle $\sigma$: this is the difference between the position of a source and the phase centre, which is where the array is pointing. For example, in \cref{fig.visibility.measure}, if we point antennas and introduce phase delays such that we are pointing the array towards Source 2, the position of Source 1 will be $\sigma$.

@@

\pg
Individual antennas of an array can be pointed mechanically, and so the loss of antenna sensitivity in the direction of interest (and therefore the loss of net interferometric array sensitivity) can be minimised. But what happens to the array itself? It is useful here to go back to the illustration of Fig. \ref{fig.arecibo.vla}. Think of each dish in the array representing a "filled" segment of a massive but empty dish. By projecting our observation in a given direction, this dish goes from circular to elliptical.
\begin{figure}[ht]
\centering
\begin{subfigure}{.40\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/lofar-uvcoverage-zenith.png}}
\caption{\label{fig.lofar.uvcoverage.zenith} $uv$-coverage of an 8-hour LOFAR observation when pointing at zenith.}
\end{subfigure}
\hfill
\begin{subfigure}{.40\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/lofar-uvcoverage-elsewhere.png}}
\caption{\label{fig.lofar.uvcoverage.elsewhere} $uv$-coverage of an 8-hour LOFAR observation when pointing 45 degrees away from zenith.}
\end{subfigure}
\caption{\label{fig.uvcoverage.lofar} Effect of array pointing on $uv$-coverage. By pointing the array 45 degrees in the $v$-axis, the array's $uv$-coverage (and thus maximum resolution) is decreased along the $v$-axis.}
\end{figure}

@@

\pg
The Point-Spread Function (PSF) of an array is thus the inverse Fourier transform of the array's $uv$-coverage. The more element in the array, the better the $uv$-coverage and thus the better the PSF. An interferometer's PSF will always be worse than an antenna dish of equivalent diameter, the UV-coverage of which is the inverse Fourier transform of a disc. This is because an interferometer behaves like a \emph{sparse} dish: the sparsity of our coverage manifests as very large point-spread function. The more baselines are available, the less sparse the UV-coverage, and so the better the PSF. This is shown in Fig. \ref{fig.vla.PSFvsUVcov}.
\begin{figure}[h!]
\centering
\resizebox{\hsize}{!}{\includegraphics{images/PSFvsUVcov.png}}
\caption{\label{fig.vla.PSFvsUVcov}Plot of the PSF associated to more and more complete UV-coverage (here, consisting of more and more elements of the VLA). Image credit: Rick Perley, 3GC4 lecture.\textcolor{red}{crop out garbage on top}}
\end{figure}

@@

\pg
We use dynamic range (DR) as a quality metric in radio images. DR is the ratio of the flux of the brightest source in the field to the flux of the faintest source in the field.  We define dynamic range as:
\begin{equation}\label{eq.DR.imag}
\mathrm{DR} = \frac{\mathrm{flux}_\mathrm{brightest}}{\mathrm{flux}_\mathrm{faintest}}
\end{equation}
The higher the DR, the deeper we image the sky, and thus the better the image. There are two big constraints to reckon with: firstly, we do not want to deconvolve the PSF from all pixels in the image, but only from those within which large amounts of underlying flux fall. Secondly, if we only deconvolve some pixels rather than performing a full deconvolution, then care must be taken not to treat artefacts in the field (which can be created from overlapping PSF sidelobes from neighbouring sources) as true flux. Doing so would lead to an attempt to deconvolve a PSF from a pixel while assuming an incorrect underlying flux value for this pixel and thus result in the introduction of further artefacts in the image.

@@

\begin{figure}[h!]
\centering
\resizebox{\hsize}{!}{\includegraphics{images/{dirty-vs-clean}.png}}
\caption{\label{fig.vla.clean} Effect of MetroCLEAN algorithm on an image of 3C295}
\end{figure}

@@

\pg
The problem of imaging radio interferometric data, of which deconvolution is but a subset, is still a very active field of research today. For more information on this topic, the NRAO website\footnote{See \href{https://www.cv.nrao.edu/~abridle/deconvol/node8.html\#SECTION00051000000000000000}{the CLEAN section of the NRAO website}.} is an excellent start for a general introduction to this topic, while \citet{2017arXiv171202078T} provides an outstanding entry point for cutting-edge work at the time of writing.

@@

\pg
Note that the weighting schemes shown here rely only on a choice between normalising visibilities by local $uv$-density vs. signal-to-noise, with Briggs weighting providing a formal method for going from one regime to another or optimising between both. Other weighting schemes exist - one can attempt to optimise data size vs decorrelation\footnote{As data is averaged, decorrelation becomes more and more important in the field. See \cref{section.RIME} for more details.}, for example \citepads[see]{2016MNRAS.462.2542A}. This can help ensure a homogeneous conditioning of the deconvolution problem in the image, by ensuring that the PSF is as peaked as possible throughout the field of interest. Part of the work of this thesis was in devising a weighting scheme dependent on calibration quality, to drastically improve deconvolution conditioning and final image quality. This weighting scheme is described in much greater detail in \cref{chapter.paper}.

\pg
We close this section by noting that data flagging \citepads[cf.]{2010ascl.soft10017O} can be considered a form of weighting scheme. This consists of assigning a null weight to those data points considered too corrupted by various instrumental or atmospheric effects (typically, Radio Frequency Interference - RFI) to be scientifically useful. Those weights which are not deemed too corrupted are assigned a unit weight. This corresponds to simply dropping those data points which would corrupt the image reconstruction by introducing unphysically strong fringes in the image.

>>>>

\pg
There are two quantities of interest to all astronomers: sensitivity and angular resolution\footnote{Other kinds of resolutions - in time and frequency, for example - are also extremely important to many astronomers, but are not what concern us here.}. An instrument's sensitivity is a function of its collecting area\footnote{It is also a function of technological factors, of course, but \emph{ceteris paribus}, a more sensitive telescope means a telescope with a wider collecting area}. Resolution, for well-designed instruments\footnote{By this, we mean that we assume that an instrument is also designed to optimise resolution.}, is limited by diffraction in the absence of atmospheric effects. This introduces specific issues in the radio domain. Radio waves have very long wavelengths - often comparable to meters, rather than the $\sim100-1000$nm wavelengths of optical light. Achieving a resolution comparable to those of optical telescopes would thus require making telescopes with apertures tens of millions of times larger than those already titanic instruments!

@@

\pg
Interferometric observations follow a very familiar pattern: the elements of the interferometric array (which can be single dishes or phased arrays\footnote{Described in \cref{sec.visibility}} of smaller antennas) measure a voltage. This voltage is correlated between antennas, as will be described shortly. Each of these correlations correspond to a single Fourier mode of the sky, as per the Zernike-van Cittert theorem (cf. \cref{sec.imag.psf}). They must be calibrated, by comparing the observations of a calibrator source and modelled visibilities. Calibration is described in much greater detail in \cref{section.RIME}. Then, the calibrated visibilities are used to reconstruct an image of the sky, which often requires deconvolution due to the sparsity of interferometric arrays.

@@

\pg
Different astrophysical sources are located very far apart, and so their signals do not interfere. They are measured additively in both antennas. Provided that the signal from both sources is coherent when observed by the dishes, the correlation between the voltages measured by antennas A and B will simply be the sum of the voltage correlations associated with individual sources. This is shown in \cref{fig.visibility}: the radio signal from two different sources, at two different locations, is picked up simultaneously by both radio antennas. Because only the correlation between both antennas is measured, the signals do not interfere. % - i.e. the interferometric signal from different sources are additive.
\pg
Note that in Fig. \ref{fig.visibility.measure}, neither source is at the zenith. This introduces the notion of the \emph{effective baseline} which is shorter than the \emph{physical baseline} when measuring a signal from a source that is not at the zenith. Sources in different positions in the sky will be measured by an interferometric array, but the same array will have different effective baselines for different sources. To point an interferometric array in a specific direction in the sky, one must introduce a phase delay in each antenna (or, for fundamental interferometer elements such as the LOFAR tiles or NenuFAR, by playing with the cable length between each dish and the correlator). This ensures that the constituent antennas' sensitivities are maximised in the direction of interest. The position of sources are then given in terms of the cardinal angle $\sigma$: this is the difference between the position of a source and the phase centre, which is where the array is pointing. For example, in \cref{fig.visibility.measure}, if we point antennas and introduce phase delays such that we are pointing the array towards Source 2, the position of Source 1 will be $\sigma$.

@@

\pg
Individual antennas of an array can be pointed mechanically, and so the loss of antenna sensitivity in the direction of interest (and therefore the loss of net interferometric array sensitivity) can be minimised. But what happens to the array itself? It is useful here to go back to the illustration of Fig. \ref{fig.arecibo.vla}. Think of each dish in the array representing a "filled" segment of a massive but empty dish. By projecting our observation in a given direction, this dish goes from circular to elliptical.
\begin{figure}[ht]
\centering
\begin{subfigure}{.40\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/lofar-uvcoverage-zenith.png}}
\caption{\label{fig.lofar.uvcoverage.zenith} $uv$-coverage of an 8-hour LOFAR observation when pointing at zenith.}
\end{subfigure}
\hfill
\begin{subfigure}{.40\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/lofar-uvcoverage-elsewhere.png}}
\caption{\label{fig.lofar.uvcoverage.elsewhere} $uv$-coverage of an 8-hour LOFAR observation when pointing 45 degrees away from zenith.}
\end{subfigure}
\caption{\label{fig.uvcoverage.lofar} Effect of array pointing on $uv$-coverage. By pointing the array 45 degrees in the $v$-axis, the array's $uv$-coverage (and thus maximum resolution) is decreased along the $v$-axis. Two colours are used, because $V_{AB}=V_{BA}$, and so only half the visibilities are stored in practice.}
\end{figure}

@@

\pg
The Point-Spread Function (PSF) of an instrument is its point source response: it will determine the angular resolution limit due to the instrument. We know that the Fourier transform of a single point source with unit flux at phase centre is a constant. If we measured the full (infinite) visibility space, then we could reconstitute a point source perfectly. In practice, however, we only ever measure a subset of the visibility space: an interferometric array's PSF is thus the inverse Fourier transform of the array's $uv$-coverage, as these are the only points for which we have information on visibility space. The more elements in the array, the better the $uv$-coverage and thus the better the PSF. An interferometer's PSF will always be worse than an antenna dish of equivalent diameter, the UV-coverage of which is the inverse Fourier transform of a disc. This is because an interferometer behaves like a \emph{sparse} dish: the sparsity of our coverage manifests as very large point-spread function. The more baselines are available, the less sparse the UV-coverage, and so the better the PSF. This is shown in Fig. \ref{fig.vla.PSFvsUVcov}.
\begin{figure}[h!]
\centering
\resizebox{\hsize}{!}{\includegraphics{images/PSFvsUVcov.png}}
\caption{\label{fig.vla.PSFvsUVcov}Plot of the PSF associated to more and more complete UV-coverage (here, consisting of more and more elements of the VLA). Image credit: Rick Perley, 3GC4 lecture.}
\end{figure}

@@

\pg
We use dynamic range (DR) as a quality metric in radio images. DR is the ratio of the flux of the brightest source in the field to the flux of the faintest source in the field.  We define dynamic range as:
\begin{equation}\label{eq.DR.imag}
\mathrm{DR} = \frac{\mathrm{flux}_\mathrm{brightest}}{\mathrm{flux}_\mathrm{faintest}}
\end{equation}
The higher the DR, the deeper we image the sky, and thus the better the image. There are two big constraints to reckon with: firstly, we do not want to deconvolve the PSF from all pixels in the image, but only from those within which large amounts of underlying flux fall. Secondly, if we only deconvolve some pixels rather than performing a full deconvolution, then care must be taken not to treat artefacts in the field (which can be created from overlapping PSF sidelobes from neighbouring sources, even with perfect calibration) as true flux. Doing so would lead to an attempt to deconvolve a PSF from a pixel while assuming an incorrect underlying flux value for this pixel and thus result in the introduction of further artefacts in the image.

@@

\begin{figure}[h!]
\centering
\resizebox{\hsize}{!}{\includegraphics{images/{dirty-vs-clean}.png}}
\caption{\label{fig.vla.clean} Effect of MetroCLEAN algorithm on a dirty image of 3C295. To the left, the dirty image; to the right, the deconvolved image (same size, resolution \& flux scale, made with the same data). As we can see, the latter is much more scientifically interesting.}
\end{figure}

@@

\pg
The problem of imaging radio interferometric data is still a very active field of research today. For more information on this topic, the NRAO website\footnote{See \href{https://www.cv.nrao.edu/~abridle/deconvol/node8.html\#SECTION00051000000000000000}{the CLEAN section of the NRAO website}.} is an excellent start for a general introduction to this topic, while \citet{2017arXiv171202078T} provides an outstanding entry point for cutting-edge work at the time of writing.

@@

\pg
%Note that the weighting schemes shown here rely only on a choice between normalising visibilities by local $uv$-density vs. signal-to-noise, with Briggs weighting providing a formal method for going from one regime to another or optimising between both. Other weighting schemes exist - one can attempt to optimise data size vs decorrelation\footnote{As data is averaged, decorrelation becomes more and more important in the field. See \cref{section.RIME} for more details.}, for example \citepads[see]{2016MNRAS.462.2542A}. This can help ensure a homogeneous conditioning of the deconvolution problem in the image, by ensuring that the PSF is as peaked as possible throughout the field of interest. Part of the work of this thesis was in devising a weighting scheme dependent on calibration quality, to drastically improve deconvolution conditioning and final image quality. This weighting scheme is described in much greater detail in \cref{chapter.paper}.

\pg
We close this section by noting that data flagging can be considered a form of weighting scheme. This consists of assigning a null weight to those data points considered too corrupted by various instrumental or atmospheric effects (typically, Radio Frequency Interference - RFI - cf. \citetads[cf.]{2010ascl.soft10017O} for an example) to be scientifically useful. Those weights which are not deemed too corrupted are assigned a unit weight. This corresponds to simply dropping those data points which would corrupt the image reconstruction by introducing unphysically strong fringes in the image.

### 2018-06-15 source/Ch2-intro/science.tex
why cheesoid exist

\section{The Low-$\nu$ Sky: Emission Mechanisms}
\pg
This section aims to bring a brief introduction to the emission mechanisms which dominate at low frequencies, and thus determine the physics accessible to extragalactic astronomers working in this band. Specifically, it will briefly describe thermal radiation detected at low frequencies (associated with dust \& protoplanetary disks, and therefore a tracer of star formation) along with free-free radiation (which is emitted from ionised plasma outflows, and thus a tracer of various physical objects e.g. young stellar objects - cf. \citet{2017ApJ...834..206C} - and starburst regions - cf. \citet{2015A&A...574A.114V}) and synchrotron radiation, which is a tracer of more violent and energetic processes (and thus associated with supernova remnants, AGN, or radio halos - cf. \citet{2010A&A...509A..68C}).

@@

\pg
Synchrotron radiation (or "magnetobremsstrahlung") occurs when the trajectory of a high-energy charged particle is deflected by a magnetic field. As the German name suggests, it is the magnetic equivalent of free-free radiation. It is the tracer of extremely violent processes, such as AGN jets. In the low-energy regime, it is also known as cyclotron radiation, after the device in which it was first measured in a laboratory. %, and in non-relativistic regimes, it is known as gyro radiation.

@@

\pg
In the optically-thick regime, we know that the most effective radiation mechanism is simple black-body radiation, and so the power spectrum at a given frequency can be defined in terms of an ``electron temperature". In the Rayleigh-Jeans approxmation (valid as we are explicitly in the low-frequency regime), this gives $S_\nu \propto \nu{5/2}$, or $\alpha=5/2$. 
In the optically-thin regime, meanwhile, we recover a mixture of the underlying power-law distribution ($S_\nu \propto e^{-\nu}$) and black-body radiation (which has a frequency-dependent $\alpha$). Empirically, we generally observe $\alpha \sim 0.7$.

\subsection{Thermal Radiation}
\pg
Also known as black-body radiation, its spectral intensity is given by Planck's law, given in Eq. \ref{eq.planck}.
\begin{equation}\label{eq.planck}
%B_\lambda (\lambda,T) = \frac{2hc^2}{\lambda^5}\left(e^{\frac{hc}{\lambda k_BT}}-1\right)^{-1}
B_\nu(\nu,T) = \frac{2h\nu^3}{c^2}\left(e^\frac{h\nu}{k_BT}-1\right)^{-1}
\end{equation}
where $B_\lambda$ is the flux density at frequency $\nu$ for a source with temperature $T$, and $k_B=1.381 \left[J/K\right]$ is the Boltzmann constant. Near protostellar disks, synchrotron emission is absorbed by the ambient interstellar medium, heating it up to an average of $\sim 10^4$ (see \citetads{2009ApJS..181..255A}, \citetads{1978ppim.book.....S} and references therein). This gives the following spectral curve:

\begin{figure*}[!h]
\centering
\includegraphics[width=0.8\textwidth]{images/ThermalEmission.png}
\caption{\label{plot.thermal}Log-Log plot of blackbody radiation emitted by a particle at $10^4$ Kelvin. Note that the y-axis is arbitrary, since it will in practice be modulated by the number of particles in a region, radiation efficiency, resolution etc.}
\end{figure*}
\pg
This is the dominant mode of emission for distant galaxies in the infrared band. In the LOFAR regime, it is not expected to dominate for extragalactic sources, but ought to be detected for resolved galaxies.

\subsection{Free-Free Radiation}
\pg
Free-free or ``bremsstrahlung" (``braking") radiation occurs when the trajectory of a charged particle is deflected by an electric field. This non-thermal emission mechanism is the dominant mechanism in HII regions (which contain ionised hydrogen), where star formation has previously taken place. It is called free-free emission because it is produced by free electrons deviating off ionised hydrogen without being captured. % as shown in Fig. \ref{plot.freefree}
%\begin{figure*}[!h]
%\centering
%\includegraphics[width=0.3\textwidth]{images/bremsstrahlung.png}
%\caption{\label{plot.freefree} Schematic showing the physical process leading to free-free emission. As we can see, the electron is scattered, but not captured. Source: \url{https://thephysicsbehind.com/2015/04/16/x-ray-tubes/}}
%\end{figure*}

>>>>

\section{The Low-$\nu$ Sky: Emission Mechanisms}
\pg
This section aims to bring a brief introduction to the emission mechanisms which dominate at low frequencies, and thus determine the physics accessible to extragalactic astronomers working in this band. Specifically, it will briefly describe
%thermal radiation detected at low frequencies (associated with dust \& protoplanetary disks, and therefore a tracer of star formation) along with 
free-free radiation (which is emitted from ionised plasma outflows, and thus a tracer of various physical objects e.g. young stellar objects - cf. \citet{2017ApJ...834..206C} and references therein - and starburst regions - cf. \citet{2015A&A...574A.114V}) and synchrotron radiation, which is a tracer of more violent and energetic processes (and thus associated with supernova remnants, AGN, or radio halos - cf. \citet{2010A&A...509A..68C}).

@@

\pg
Synchrotron radiation (or "magnetobremsstrahlung") occurs when the trajectory of a relativistic charged particle is deflected by a magnetic field. As the German name suggests, it is the magnetic equivalent of free-free radiation. It is the tracer of very energetic processes, such as the interaction between AGN jets and the intergalactic medium. In the low-energy regime, it is also known as cyclotron radiation, after the device in which it was first measured in a laboratory. %, and in non-relativistic regimes, it is known as gyro radiation.

@@

\pg
In the optically-thick regime, we know that the most efficient radiation mechanism is simple black-body radiation, and so the power spectrum at a given frequency can be defined in terms of an ``electron temperature". In the Rayleigh-Jeans approxmation (valid as we are explicitly in the low-frequency regime), this gives $S_\nu \propto \nu{5/2}$, or $\alpha=5/2$. 
In the optically-thin regime, meanwhile, we recover a mixture of the underlying power-law distribution ($S_\nu \propto e^{-\nu}$) and black-body radiation (which has a frequency-dependent $\alpha$). Empirically, we generally observe $\alpha \sim 0.7$.

%\subsection{Thermal Radiation}
%\pg
%Also known as black-body radiation, its spectral intensity is given by Planck's law, given in Eq. \ref{eq.planck}.
%\begin{equation}\label{eq.planck}
%%B_\lambda (\lambda,T) = \frac{2hc^2}{\lambda^5}\left(e^{\frac{hc}{\lambda k_BT}}-1\right)^{-1}
%B_\nu(\nu,T) = \frac{2h\nu^3}{c^2}\left(e^\frac{h\nu}{k_BT}-1\right)^{-1}
%\end{equation}
%where $B_\lambda$ is the flux density at frequency $\nu$ for a source with temperature $T$, and $k_B=1.381 \left[J/K\right]$ is the Boltzmann constant. Near protostellar disks, synchrotron emission is absorbed by the ambient interstellar medium, heating it up to an average of $\sim 10^4$ (see \citetads{2009ApJS..181..255A}, \citetads{1978ppim.book.....S} and references therein). Assuming that the interstellar medium is optically thick at all frequencies, this gives the following spectral curve:
%
%\begin{figure*}[!h]
%\centering
%\includegraphics[width=0.8\textwidth]{images/ThermalEmission.png}
%\caption{\label{plot.thermal}Log-Log plot of blackbody radiation emitted by a particle at $10^4$ Kelvin. Note that the y-axis is arbitrary, since it will in practice be modulated by density, radiation efficiency, resolution etc.}
%\end{figure*}
%\pg
%This is the dominant mode of emission for distant galaxies in the infrared band, but is not expected to contribute in the LOFAR regime for extragalactic sources. It ought to be detected for nearby resolved galaxies, however.

\subsection{Free-Free Radiation}
\pg
Free-free or ``bremsstrahlung" (``braking") radiation occurs when the trajectory of a charged particle is deflected by an electric field. This non-thermal emission mechanism is the dominant mechanism in HII regions (which contain ionised hydrogen), where star formation has previously taken place. It is called free-free emission because it is produced by free electrons deviating off ionised hydrogen without being captured. When the underlying electron distribution is Maxwellian, it is referred to as thermal Bremsstrahlung.% as shown in Fig. \ref{plot.freefree}
%\begin{figure*}[!h]
%\centering
%\includegraphics[width=0.3\textwidth]{images/bremsstrahlung.png}
%\caption{\label{plot.freefree} Schematic showing the physical process leading to free-free emission. As we can see, the electron is scattered, but not captured. Source: \url{https://thephysicsbehind.com/2015/04/16/x-ray-tubes/}}
%\end{figure*}

### 2018-06-15 source/Ch3-RIME/RIME.tex
why cheesoid exist

\pg
Here, effects such as our radio beam lie in the Jones matrices. By performing the analysis shown in Sect. 3.1 of \citetads{2001isra.book.....T}, we can rewrite this expression in terms of a sine projection of the sphere onto the $(l,m)$ plane tangential at field centre. The integral then becomes:
\begin{equation}
\Vis_{pq} = \int_{l,m} \Jones_p(\direction) \Bmatrix(\direction) \Jones_q^H(\direction)\frac{dl dm}{n}
\end{equation}

@@

\pg
Earlier calibration models and algorithm rely on the hypothesis that all baselines see the same sky. This is the \emph{fundamental premise of self-calibration}. The topic of self-cal is covered in \cref{section.calibration.2gc}. This premise only holds when all Direction-Dependent Effects (henceforth referred to as DDEs) are identical for all baselines (and therefore for all antennas). This is a necessary condition for the apparent sky $\Bmatrix_{pq}$ to be the same for all baselines\footnote{$\Bmatrix_{pq}$ would then correspond to the `true' sky attenuated by the power beam, in the traditional view}, as it allows us to write:
\begin{align}
\Ejones_p(\lvec) \Wjones_p(\lvec)&= \Ejones(\lvec) \Wjones(\lvec)\\
\Bmatrix_{pq}(\lvec) = \Bmatrix_{app}(\lvec) &= \Ejones(\lvec) \Wjones(\lvec) \Bmatrix)(\lvec) \Wjones^H(\lvec) \Ejones^H(\lvec)
\end{align}

>>>>

\pg
Here, effects such as our array's beam lie in the Jones matrices. By performing the analysis shown in Sect. 3.1 of \citetads{2001isra.book.....T}, we can rewrite this expression in terms of a sine projection of the sphere onto the $(l,m)$ plane tangential at field centre. The integral then becomes:
\begin{equation}
\Vis_{pq} = \int_{l,m} \Jones_p(\direction) \Bmatrix(\direction) \Jones_q^H(\direction)\frac{dl dm}{n}
\end{equation}

@@

\pg
Earlier calibration models and algorithm rely on the hypothesis that all baselines see the same sky. This is the \emph{fundamental premise of self-calibration}. The topic of self-cal is covered in \cref{section.calibration.2gc}. This premise only holds when all Direction-Dependent Effects (henceforth referred to as DDEs) are identical for all baselines (and therefore for all antennas). This is a necessary condition for the apparent sky $\Bmatrix_{pq}$ to be the same for all baselines\footnote{$\Bmatrix_{pq}$ would then correspond to the `true' sky attenuated by the ``power beam", which corresponds to our instrument's directional sensitivity in the traditional view}, as it allows us to write:
\begin{align}
\Ejones_p(\lvec) \Wjones_p(\lvec)&= \Ejones(\lvec) \Wjones(\lvec)\\
\Bmatrix_{pq}(\lvec) = \Bmatrix_{app}(\lvec) &= \Ejones(\lvec) \Wjones(\lvec) \Bmatrix)(\lvec) \Wjones^H(\lvec) \Ejones^H(\lvec)
\end{align}
